---
title: "Seminar Code - Markdown"
author: "Lukas Pitz"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Cleaning and Adjustment

The code in this part of the Markdown is used to load and clean the data set provided by Nicolas Carbone on his kaagle account.
To run this code, the path on your local device needs to get adjusted. See vaiable "my_path" below.

IMPORTANT: This part may be skipped, we provide a cleaned version of the data set in the zip-file. 

## Packages 

```{r}
# Used Packages
library(tidyverse) # data cleaning and manipulation
library(dplyr) # data cleaning and manipulation
library(Hmisc) # Statistical parameters
library(glmnet) # Lasso, Ridge and elastic Net (lambda via cross-validation)
library(hdm) # Lasso by Chernozhukov, Hansen and Spindler (lambda via theoretical foundation)
library(mosaic)  # standardizing variables
library(lmtest) # tests for linear regression model
library(nlme) # GLS
library(olsrr) # Variance Inflation Factor (VIF)
library(rpart) # Regression Tree models
library(rpart.plot) # Visualization tool for Regression Tree models
library(ipred) # Regression Tree models with bagging
library(randomForest) # Random Forest models
library(ranger) # Alternative package for Random Forest models
library(networkD3) # Create Sankey diagrams
library(latex2exp) # enables latex expressions in plots
library(aod) # Implementation of Wald tests
library(gbm) # Gradient Boosted Tree models
library(xgboost) # Faster implementation of Gradient Boosted Tree models

# Path to data on my own device
my_path = 'G:/Meine Ablage/Uni Köln/Seminar Empirical Methods and Data Analysis/Data/'

```

## Loading and transforming data

```{r}
# Data set from 2014
df_2014 = read.csv(paste(my_path,'2014_Financial_Data.csv', sep='')) %>%
    add_column(year = 2014) %>% # add column with year
    rename('price_var' = 'X2015.PRICE.VAR....') %>% # rename outcome variable
    relocate('year', .before = 'X') %>% # relocate year column
    relocate('price_var', .after = 'X') %>% # relocate outcome variable column
    arrange(X) # arrange rows by company name

# Data set from 2015
df_2015 = read.csv(paste(my_path,'2015_Financial_Data.csv', sep='')) %>%
    add_column(year = 2015) %>%
    rename('price_var' = 'X2016.PRICE.VAR....') %>%
    relocate('year', .before = 'X') %>%
    relocate('price_var', .after = 'X') %>%
    arrange(X)

# Data set from 2016
df_2016 = read.csv(paste(my_path,'2016_Financial_Data.csv', sep='')) %>%
    add_column(year = 2016) %>%
    rename('price_var' = 'X2017.PRICE.VAR....') %>%
    relocate('year', .before = 'X') %>%
    relocate('price_var', .after = 'X') %>%
    arrange(X)

# Data set from 2017
df_2017 = read.csv(paste(my_path,'2017_Financial_Data.csv', sep='')) %>%
    add_column(year = 2017) %>%
    rename('price_var' = 'X2018.PRICE.VAR....') %>%
    relocate('year', .before = 'X') %>%
    relocate('price_var', .after = 'X') %>%
    arrange(X)

# Data set from 2018
df_2018 = read.csv(paste(my_path,'2018_Financial_Data.csv', sep='')) %>%
    add_column(year = 2018) %>%
    rename('price_var' = 'X2019.PRICE.VAR....') %>%
    relocate('year', .before = 'X') %>%
    relocate('price_var', .after = 'X') %>%
    arrange(X)

# Find stocks by name which are included in each data frame 
common_stocks = Reduce(intersect, list(df_2014$X,df_2015$X,df_2016$X,df_2017$X,df_2018$X))

# Delete stocks which are not included in each data frame
df_2014 = df_2014[df_2014$X %in% common_stocks,]
df_2015 = df_2015[df_2015$X %in% common_stocks,]
df_2016 = df_2016[df_2016$X %in% common_stocks,]
df_2017 = df_2017[df_2017$X %in% common_stocks,]
df_2018 = df_2018[df_2018$X %in% common_stocks,]

# Merge data frames
df_full = rbind(df_2014,df_2015,df_2016,df_2017,df_2018) %>%
          arrange(X)
```

## Data quality check A (before cleaning)

Get a first overview of data quality and data types. 

```{r}
# Check data types
table(sapply(df_full, class))

# Number of columns with more than 30%  NA's
sum(sapply(df_full, function(x){sum(is.na(x))/nrow(df_full)}) > 0.3)

# Number of columns with more than 30% zeros (replace NA's with zeros because R has issues with handling NA's)
df_full_no_NA = replace(df_full, is.na(df_full),0)

# Count the number of columns with more than 30% zeros in them
sum(sapply(df_full_no_NA,function(x){sum(x==0)/nrow(df_full_no_NA)}) > 0.3) 
```

## Manuel Data Cleaning and Manipulation: Zeros and dummy variables

```{r}
### Excluding columns which contain many missing observations (NA's and zeros) ###

# Delete class column (contains non relevant information)
df_full_1 = df_full %>% subset(select = -c(Class))

# Create new data frame which contains only columns with less than 30% zeros as observations
df_full_1 = df_full_no_NA[sapply(df_full_no_NA,function(x){sum(x==0)/nrow(df_full_no_NA)}) < 0.3]

### Creating sector dummy variables ###
df_full_1$Basic_Materials         = ifelse(df_full$Sector == 'Basic Materials',1,0)
df_full_1$Communication_Services  = ifelse(df_full$Sector == 'Communication Services',1,0)
df_full_1$Consumer_Cyclical       = ifelse(df_full$Sector == 'Consumer Cyclical',1,0)
df_full_1$Consumer_Defensive      = ifelse(df_full$Sector == 'Consumer Defensive',1,0)
df_full_1$Energy                  = ifelse(df_full$Sector == 'Energy',1,0)
df_full_1$Financial_Services      = ifelse(df_full$Sector == 'Financial Services',1,0)
df_full_1$Healthcare              = ifelse(df_full$Sector == 'Healthcare',1,0)
df_full_1$Industrials             = ifelse(df_full$Sector == 'Industrials',1,0)
df_full_1$Real_Estate             = ifelse(df_full$Sector == 'Real Estate',1,0)
df_full_1$Technology              = ifelse(df_full$Sector == 'Technology',1,0)
df_full_1$Utilities               = ifelse(df_full$Sector == 'Utilities',1,0)

# Delete sector column
df_full_1 = df_full_1 %>% subset( select = -c(Sector))

```

## Manuel Data Cleaning and Manipulation: Delete identical columns

```{r}
# Number of identical pairs of columns
sum(duplicated(as.list(df_full_1)))

# Names of column pairs
colnames(df_full_1[duplicated(as.list(df_full_1))])

# Drop one of the duplicated columns
df_full_1 = df_full_1[!duplicated(as.list(df_full_1))]

```

## Manuel Data Cleaning and Manipulation: Outliers

We experimented with two approaches to identify outliers. Firstly by looking at the standard deviation of each data point from its column mean. Secondly, by declaring data points in the 0.01% and 99.9% quantile (columnwise) as outliers. We settled with the second approach, because the quantiles are not affected by the magnitude of the outliers their self, compared to the standard deviation which proved to be useful.  

```{r}

### Solution A: standard deviations from the means
# 
# # Determine standard deviation per column
# sd_df_full = df_full_1 %>%
#   apply(2,sd) %>%
#   as.data.frame() %>%
#   rownames_to_column()
# 
# ### Imputing rows with outliers ###
# df_full_2 = df_full_1
# 
# # Step one: Replace all outliers with NAs
# 
# # loop over columns (start at i=3 to skip year and name columns; skip the last 11 dummy variable columns)
# for (i in 3:(length(df_full_2) - 11)){ 
#   
#   for(j in 1:nrow(df_full_2)){ # loop over rows
#    
#      if(!((df_full_2[j,i] < (mean(df_full_1[,i]) + 5 * sd_df_full[i,2])) &
#        (df_full_2[j,i] > (mean(df_full_1[,i]) - 5 * sd_df_full[i,2])))){
#         df_full_2[j,i] = NA}
#   }
#   }
# 
# # Step two: Impute all NAs with the column mean 
# df_full_2 = df_full_2 %>% mutate_all(~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x))
# 
# #Compare Results
# summary(df_full_1[,'price_var']) # before imputing outliers
# summary(df_full_2[,'price_var']) # after imputing outliers


### Solution B: Trimming the 0.01% and 99.9% quantile per column ###

df_full_2 = df_full_1

# Determine quantile thresholds per column

quant_df_full = df_full_1 %>%
  subset(select= -c(year,X)) %>%
  apply(2, quantile, probs = c(0.001,0.999)) %>%
  as.data.frame() 


# Step 1: Replace all outliers with NAs

for (i in 3:(length(df_full_2) - 11)){ # loop over columns
  
  for(j in 1:nrow(df_full_2)){ # loop over rows
   
     if(
       df_full_2[j,i] < quant_df_full[1,i-2] || # smaller than the lower 0.1% quantile threshold
       df_full_2[j,i] > quant_df_full[2,i-2] # larger than the 99.9% quantile threshold
       ){df_full_2[j,i] = NA}
  }
  }

# Step 2: Impute all NAs with the column mean 
df_full_2 = df_full_2 %>% mutate_all(~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x))

#Compare Results
summary(df_full_1[,'price_var']) # before imputing outliers
summary(df_full_2[,'price_var']) # after imputing outliers

```

## Manuel Data Cleaning and Manipulation: Multicollinearity

We utilize the Variance Inflation Factor (VIF) to detect and exclude features within our high-dimensional data set which contribute little additional information due to their severe collinearity to other considered features. For more details see Appenix A in our term paper.


```{r}
# Delete one sector dummy variable to prevent perfect collinearity
df_full_3 = df_full_2 %>% subset(select = -c(year,X,Utilities))

# Check for multicollinearity via VIF (Variance Inflation Factor)

# 1 Step: Run without loop
test_OLS = lm(data=df_full_3,formula= price_var ~ . )

vif = ols_vif_tol(test_OLS)

vif = vif[vif$VIF > 10,] %>% arrange(desc(VIF)) # select all features with VIF>10 and arrange in desc order

print(nrow(vif)) # number of variables with a vif higher than 10 and therefore evidence for multicollinearity 

df_full_3 = df_full_3[ , -which(names(df_full_3) %in% c(vif[1,1]))] # drop column with highest VIF

# 2. Step: Run with loop
while (nrow(vif) > 2 ){
test_OLS = lm(data=df_full_3,formula= price_var ~ . )

vif = ols_vif_tol(test_OLS)

vif = vif[vif$VIF > 10,] %>% arrange(desc(VIF)) # select all features with VIF>10 and arrange in desc order

df_full_3 = df_full_3[ , -which(names(df_full_3) %in% c(vif[1,1]))] # drop column with highest VIF

print(nrow(vif))
}

# Add year and company name column back to data frame
df_full_3 = cbind(df_full_2$year, df_full_2$X, df_full_3) %>%
            rename('year' = 'df_full_2$year') %>%
            rename('X' = 'df_full_2$X')

# Number of columns drooped due to multicollinearity
ncol(df_full_2) - ncol(df_full_3)

# List of variables which got dropped due to multicollinearity
setdiff(union(names(df_full_2), names(df_full_3)), intersect(names(df_full_2), names(df_full_3)))
```

## Data quality check B (after cleaning)

Scan data after cleaning for irregularities which could hint to errors in the code or further need for cleaning. 

```{r}
### Check if cleaning worked ###

# Number of columns with NA's
sum(sapply(df_full_3, function(x){sum(is.na(x))/nrow(df_full_2)}) > 0)

# Number of columns with more than 30%  zeros (should expect 11 due to the created dummy variables)
sum(sapply(df_full_3,function(x){sum(x==0)/nrow(df_full_2)}) > 0.3)

# Check number of companies in data set after cleaning
length(unique(df_full_3$X))

# Check if columns only contain one value
sum(sapply(df_full_3,function(x) length(unique(x))==1))

# Check data types
table(sapply(df_full_3, class))

### Check if outlier imputation worked properly ###

# Solution A applied:

# # Determine distance of each observation to the pre imputation mean in units of standard deviations
# outliers = df_full_3

# for (i in 3:(length(df_full_3) - 11)){ # start at i=3 due to the character column
#   outliers[,i] = abs(df_full_3[,i] - mean(df_full_1[,i])) / sd_df_full[i,2]
# }
 
# # Results should be in [0,5]
# max(outliers[,-c(1,2)])
# min(outliers[,-c(1,2)])

# Solution B applied:

```


## Data checkpoint: Load cleaned data set

Possibility to load data set incorporating cleaned data to avoid re-run of code above. 

```{r}
#write.csv(df_full_3,paste(my_path,'cleaned_full_data_set.csv', sep=''),row.names=FALSE)

# Read in cleaned data set
df_full_3 = read.csv(paste(my_path,'cleaned_full_data_set.csv', sep=''))
  
```

## Split data set in training and test data set

```{r}
# Training data set (2014-2017)
df_train = df_full_3 %>% filter(year %in% c(2014,2015,2016,2017))

# Test data set (2018)
df_test = df_full_3 %>% filter(year == 2018)

# Function for Root-Mean-Squared-Error (RMSE)
RMSE = function(y, yhat){
  return(sqrt( mean( (y - yhat)^2) ) )}
```


# Machine Learn (ML) Models

The following code chunks correspond to different ML models, we apply. More details w.r.t. each models theoretical concept and our applied tuning process are provided in our term paper within the corresponding (sub-)sections.


## Lasso Model

Lasso model implementation by package: glmnet

```{r}
### Hyperperameter tuning for Lasso model ###
 
# Independent  regressors
x_train = data.matrix(subset(df_train, select = -c(year,X,price_var)))
x_test = data.matrix(subset(df_test, select = -c(year,X,price_var)))

# Dependent variable
y_train = as.vector(df_train$price_var) 
y_test = as.vector(df_test$price_var) 

# Hyper parameter tuning (lambda)
set.seed(123)

find_best_lambda_lasso = cv.glmnet(x=x_train,y=y_train,
                        type.measure = 'mse', # loss measure
                        nfolds = 10, # number of folds for cross-validation
                        alpha = 1, # alpha = 1 corresponds to the Lasso-estimator
                        standardize = TRUE, # normalize variables
                        intercept = TRUE,
                        trace.it=TRUE)

print(find_best_lambda_lasso)

# Optimal lambda value that minimizes training MSE
best_lambda_lasso = find_best_lambda_lasso$lambda.min

lowest_MSE_lasso =  min(find_best_lambda_lasso$cvm)

plot(find_best_lambda_lasso) # Plot MSE over lambda values


### Training Lasso model using best lambda ###

lasso_model_glmnet = glmnet(x=x_train,y=y_train,
                        lambda = best_lambda_lasso,
                        type.measure = 'mse', # loss measure
                        alpha = 1, # alpha = 1 corresponds to the Lasso-estimator
                        standardize = TRUE, # normalize variables
                        intercept = TRUE,
                        trace.it=FALSE)

coef_exact_lasso = coef(lasso_model_glmnet, s = best_lambda_lasso, exact=T, x=x_train,y=y_train) %>%as.vector()
sum(coef_exact_lasso !=0) # count the non-zero coefficients


### Testing model on in and out-of-sample data ###

# Fit the best selected model on the test data
lasso_train_fitted = cbind(1,x_train) %*% coef_exact_lasso
lasso_test_fitted = cbind(1,x_test) %*% coef_exact_lasso

RMSE_Lasso_train = RMSE(df_train$price_var,lasso_train_fitted)
RMSE_Lasso_test = RMSE(df_test$price_var,lasso_test_fitted)
```


## Ridge Approach

Ridge model implementation by package: glmnet

```{r}
### Hyperperameter tuning for Ridge model ###
 
# Independent  regressors
x_train = data.matrix(subset(df_train, select = -c(year,X,price_var)))
x_test = data.matrix(subset(df_test, select = -c(year,X,price_var)))

# Dependent variable
y_train = as.vector(df_train$price_var) 
y_test = as.vector(df_test$price_var) 

# Hyper parameter tuning (lambda)
set.seed(123)

find_best_lambda_ridge = cv.glmnet(x=x_train,y=y_train,
                        type.measure = 'mse', # loss measure
                        nfolds = 10, # number of folds for cross-validation
                        alpha = 0, # alpha = 0 corresponds to the ridge-estimator
                        standardize = TRUE, # normalize variables
                        intercept = TRUE,
                        trace.it=TRUE)

print(find_best_lambda_ridge)

# Find optimal lambda value that minimizes MSE
best_lambda_ridge = find_best_lambda_ridge$lambda.min

lowest_MSE_ridge =  min(find_best_lambda_ridge$cvm)

plot(find_best_lambda_ridge) # Plot MSE over lambda values


### Training Lasso model using best lambda ###

ridge_model_glmnet = glmnet(x=x_train,y=y_train,
                        lambda = best_lambda_ridge,
                        type.measure = 'mse', # loss measure
                        alpha = 0, # alpha = 1 corresponds to the ridge-estimator
                        standardize = TRUE, # normalize variables
                        intercept = TRUE,
                        trace.it=FALSE)

coef_exact_ridge = coef(ridge_model_glmnet, s = best_lambda_ridge, exact=T, x=x,y=y) %>% as.vector()

sum(coef_exact_ridge !=0) # count the non-zero coefficients (with ridge should be 170)


### Testing model on in and out-of-sample data ###

# Fit the best selected model on the test data
ridge_train_fitted = cbind(1,x_train) %*% coef_exact_ridge
ridge_test_fitted = cbind(1,x_test) %*% coef_exact_ridge

RMSE_Ridge_train = RMSE(df_train$price_var,ridge_train_fitted)
RMSE_Ridge_test = RMSE(df_test$price_var,ridge_test_fitted)
```


## Elastic Net Approach

Elastic Net model implementation by package: glmnet

```{r}
### Hyperperameter tuning for Elastic Net model ###
 
# Independent  regressors
x_train = data.matrix(subset(df_train, select = -c(year,X,price_var)))
x_test = data.matrix(subset(df_test, select = -c(year,X,price_var)))

# Dependent variable
y_train = as.vector(df_train$price_var) 
y_test = as.vector(df_test$price_var) 

# Hyper parameter tuning (lambda and alphas)

# Tuning elastic net to find best alpha and lambda combination (model with lowest MSE across all alphas)

alphas = seq(from=0,to=1, length.out=100) # create vector with all alphas desired to test

best_lamda_elastic = c() # create empty vector to store best lambda per run 

Lowest_MSE_elastic = c() # create empty vector to store lowest MSE per run 


for (i in 1:length(alphas)){
set.seed(123) # reproducibility

find_best_lambda_alpha_elastic = cv.glmnet(x=x_train,y=y_train,
                        type.measure = 'mse', # loss measure 
                        nfolds = 10, # number of folds for cross-validation
                        alpha = alphas[i],
                        standardize = TRUE, # normalize variables
                        intercept = TRUE,
                        trace.it = FALSE) # no need for progress bars

best_lamda_elastic[i] = find_best_lambda_alpha_elastic$lambda.min
Lowest_MSE_elastic[i] = min(find_best_lambda_alpha_elastic$cvm)
}

# Store results in data frame and sort depending on Lowest_MSE_elastic ascending 
elastic_results = cbind(alphas,best_lamda_elastic,Lowest_MSE_elastic) %>%
                  as.data.frame()

elastic_results %>% arrange(Lowest_MSE_elastic)

# Visualization
ggplot(elastic_results, aes(x=alphas,y=Lowest_MSE_elastic)) +
  geom_point() +
  geom_line() +
  labs(x="alpha",y="Validation MSE") +
  theme_light()

```


## Regression Tree model

Regression Tree model implementation by package: rpart 

```{r}
### Hyperparameter tuning for Regression Tree model ###

# Create grid of hyperparameters for tuning 
hyper_grid = expand.grid(
  minsplit = seq(5, 20, 1), # min obs. in a nod to perform split
  maxdepth = seq(1, 10, 1)) # max depth of the final tree

nrow(hyper_grid) # number of models to be specified

models = list() # create empty list to store tree models in 

# Loop over the different hyperparameters
for (i in 1:nrow(hyper_grid)) {
  
  # get minsplit and maxdepth values at row i
  minsplit = hyper_grid[i,1]
  maxdepth = hyper_grid[i,2]
  
  set.seed(123)

  # train a model and store in the list
  models[[i]] = rpart(
    formula = price_var ~ .,
    data    = df_train %>% subset(select = -c(year,X)),
    method  = "anova",
    control = list( minsplit = minsplit, # min obs. in a nod to perform split
                    minbucket = round(minsplit/3), # min number of obs. in terminal leaf
                    xval = 10, # number of folds in cross-validation
                    maxdepth = maxdepth) # max depth of the final tree
  )
  print(i) # to check progress of loop
  }

# # Function to get cp
 get_cp = function(x) {
   min    = which.min(x$cptable[, "xerror"])
   cp = x$cptable[min, "CP"]}

# Function to get minimum error across models
get_min_error = function(x) {
  min    = which.min(x$cptable[, "xerror"])
  xerror = x$cptable[min, "xerror"]}

# Add corresponding hyperparameters to grid
hyper_grid = hyper_grid %>%
             mutate(cp = purrr::map_dbl(models, get_cp),
                    error = purrr::map_dbl(models, get_min_error)) %>%
             arrange(error) 

hyper_grid


### Training Regression Tree model using optimal hyperparameters ###

set.seed(123)

tree_model= rpart(
    formula = price_var ~ .,
    data    = df_train %>% subset(select = -c(year,X)),
    method  = "anova",
    control = list( minsplit = hyper_grid[1,1], # min obs. in a nod to perform split
                    minbucket = round(hyper_grid[1,1]/3), # min number of obs. in terminal leaf
                    xval = 10, # number of folds in cross-validation
                    cp = hyper_grid[1,3], # complexity parameter
                    maxdepth = hyper_grid[1,2])) # max depth of the final tree

# Values for split process
printcp(tree_model)

# Visualize Tree
rpart.plot(tree_model, type = 2)


### Testing model on in and out-of-sample data ###
regression_tree_fitted_train = predict(tree_model,
                                 newdata = df_train %>% subset(select = -c(year,X)) )

regression_tree_fitted_test = predict(tree_model,
                                 newdata = df_test %>% subset(select = -c(year,X)) )

RMSE_Tree_train = RMSE(df_train$price_var,regression_tree_fitted_train)
RMSE_Tree_test = RMSE(df_test$price_var,regression_tree_fitted_test)
```


## Regression Tree model utilizing bagging

Regression Tree model utilizing bagging implementation by package: ipred 

```{r}
### Hyperparameter tuning for Regression Tree model utilizing bagging ###

# Sequence of number of trees 
ntree = seq(from=10,to=100,by=2)

sum(ntree) # number of total trees to be estimated 

# Create empty vector to store out-of-bag RMSE values
rmse = vector(mode = "numeric", length = length(ntree))

# Bagged Regression Tree model
for (i in 1:length(ntree)) {
  
  set.seed(123) # Reproducibility 
  
  model = bagging(
  formula = price_var ~ .,
  data    = df_train %>% subset(select = -c(year,X)),
  coob    = TRUE, # use the out-of-bag sample to calculate validation error
  nbagg   = ntree[i])
  
  rmse[i] = model$err # store the out-of-bag validation RMSE in vector
  
  print(i) # to check progress of loop

  }

# Data Frame of number of trees with corresponding RMSE
rmse_ntree = as.data.frame(cbind(ntree,rmse)) %>% arrange(rmse)

# Plot out-of-bag RMSE vs. number of trees
ggplot(data = data.frame(ntree,rmse), aes(x=ntree,y=rmse)) +
  geom_line(linewidth=1) +
  geom_vline(xintercept = rmse_ntree[1,1], color="#eb8060",lty = "dashed",linewidth=1) + # number of tree with lowest out-of-bag RMSE
  labs(x="Number of Trees",y="OOB RMSE") + 
  theme_classic()


### Training Regression Tree model utilizing bagging using optimal hyperparameters ###

set.seed(123) # Reproducibility

bag_tree_model = bagging(
                        formula = price_var ~ .,
                        data    = df_train %>% subset(select = -c(year,X)),
                        coob    = TRUE, # use the out-of-bag sample to calculate validation error
                        nbagg   = rmse_ntree[1,1])


### Testing model on in and out-of-sample data ###

regression_bag_tree_fitted_train = predict(bag_tree_model,
                                 newdata = df_train %>% subset(select = -c(year,X)) )

regression_bag_tree_fitted_test = predict(bag_tree_model,
                                 newdata = df_test %>% subset(select = -c(year,X)) )

RMSE_Bag_Tree_train = RMSE(df_train$price_var,regression_bag_tree_fitted_train)
RMSE_Bag_Tree_test = RMSE(df_test$price_var,regression_bag_tree_fitted_test)
```


## Random Forest (RF) model

Random Forest model utilizing bagging implementation by packages: randomForest and ranger

```{r}
### Hyperperameter tuning for Random Forest model ###


### Step 1: Find optimal number of trees (using RF implementation by package: randomForest) ###

set.seed(123) # Set set to make results reproducible

# Estimate RF with up to 1000 trees
RFmodel_tree = randomForest(price_var ~ ., 
                       data = df_train %>% subset(select = -c(year,X)),
                       ntree = 1000, # max. number of trees averaged over
                       importance = TRUE)


# Extract out-of-bag MSE and number of trees
RFmodel_OOB_MSE_ntree = data.frame(cbind(1:1000,RFmodel_tree$mse))

# Visualize error rate while averaging over more and more trees 
ggplot(data = RFmodel_OOB_MSE_ntree, aes(x=X1,y=X2)) +
  geom_line(linewidth=1) + 
  geom_vline(xintercept = which.min(RFmodel_OOB_MSE_ntree$X2), color="#eb8060",lty = "dashed",linewidth=1) + # number of tree with lowest out-of-bag RMSE
  labs(x="Number of Trees",y="OOB MSE") + 
  theme_classic()


### Step 2: Find the optimal values for: (using RF implementation by package: ranger) ###

# mtry =  max number of randomly drawn split variables to consider at each node 
# min.node.size = minimal number of observations in a (potential node) to consider a split
# min.bucket = minimal observations in the terminal leaf
# max.depth = maximal depth of tree models (in terms of levels)

default_mtry = floor(sqrt(length(df_train %>% subset(select = -c(year,X))))) # default mtry chose by rule of thumb 

# Create grid of hyperparameters for tuning 
hyper_grid_RF = expand.grid(
  min.node.size = seq(1, 15, 2), 
  maxdepth = seq(1, 10, 2), 
  mtry = seq(2, 2*default_mtry - 2, by = 1))  
  
nrow(hyper_grid_RF) # number of RF models to be trained

OOB_MSE_RF_base_learner = c()

set.seed(123) # Reproducibility

for (i in 1:nrow(hyper_grid_RF)) {

RFmodel_base_learner = ranger(
                  formula   = price_var ~ ., 
                  data      = df_train %>% subset(select = -c(year,X)), 
                  num.trees = 500,
                  mtry      = hyper_grid_RF[i,3],
                  min.node.size = hyper_grid_RF[i,1],
                  min.bucket = round(hyper_grid_RF[i,1]/3),
                  max.depth = hyper_grid_RF[2,i],
                  oob.error= TRUE,
                  verbose = FALSE) # Turn of internal progress report 

OOB_MSE_RF_base_learner[i] = RFmodel_base_learner$prediction.error

print(round((i/nrow(hyper_grid_RF)*100),4)) # to check progress of loop in percent
}

RFmodel_OOB_MSE_base_learner = cbind(hyper_grid_RF,OOB_MSE_RF_base_learner) %>% # add OOB MSE to hyper grid
  arrange(OOB_MSE_RF_base_learner)


### Fitting model using the optimal tuning parameters ### 

set.seed(123) # Reproducibility

Random_Forest_model = ranger(
                        formula   = price_var ~ ., 
                        data      = df_train %>% subset(select = -c(year,X)), 
                        num.trees = 500, 
                        mtry      = RFmodel_OOB_MSE_base_learner[3,1],
                        min.node.size = RFmodel_OOB_MSE_base_learner[1,1],
                        min.bucket = round(RFmodel_OOB_MSE_base_learner[1,1]/3),
                        max.depth = RFmodel_OOB_MSE_base_learner[2,1],
                        oob.error= TRUE,
                        verbose = FALSE) # Turn of internal progress report


### Testing model on in and out-of-sample data ###

regression_random_forest_fitted_train = predict(Random_Forest_model,
                                          data = df_train %>% subset(select = -c(year,X)))

regression_random_forest_fitted_test = predict(Random_Forest_model,
                                          data = df_test %>% subset(select = -c(year,X)))

RMSE_Random_Forest_train = RMSE(df_train$price_var,regression_random_forest_fitted_train$predictions)
RMSE_Random_Forest_test = RMSE(df_test$price_var,regression_random_forest_fitted_test$predictions)
```


## Gradient Boosted Regression Trees (GBRT) 

Gradient Boosted Regression Trees model implementation by package: xgboost

```{r}
### Hyperperameter tuning for Gradient Boosting Regression Tree Model ###

# Independent  regressors
x_train = data.matrix(subset(df_train, select = -c(year,X,price_var)))
x_test = data.matrix(subset(df_test, select = -c(year,X,price_var)))

# Dependent variable
y_train = data.matrix(df_train$price_var) 
y_test = data.matrix(df_test$price_var) 


# Set grid search values for tuning
hyper_grid_GBRT = expand.grid(
  eta = c(0.005,0.01, 0.05, 0.1), # learning rate for Gradient Descent Algorithm  
  max_depth = c(5, 7, 9, 11), # maximum tree depth 
  min_child_weight = c(5, 7, 9, 11), # minimum number of observations required in terminal node
  subsample = c(0.65, 0.8, 1), # fraction of training data used to train each tree
  colsample_bytree = c(0.7,0.8, 0.9, 1), # fraction of features used to train each tree
  optimal_trees = 0,               # place to store later results 
  min_RMSE = 0)                     # place to store later results 


nrow(hyper_grid_GBRT) # number of different models to train


for(i in 1:nrow(hyper_grid_GBRT)) {
  
# Create temporary hyperparameter list
params = list(
    eta = hyper_grid_GBRT$eta[i],
    max_depth = hyper_grid_GBRT$max_depth[i],
    min_child_weight = hyper_grid_GBRT$min_child_weight[i],
    subsample = hyper_grid_GBRT$subsample[i],
    colsample_bytree = hyper_grid_GBRT$colsample_bytree[i])
  
set.seed(123) # for reproducibility
  
GBRT_tuning = xgb.cv(
    params = params,
    data = x_train,
    label = y_train,
    metrics = "rmse", # use RMSE for validation in cross-validation process
    nrounds = 1000, # Max number of iterations (different tree models)
    nfold = 5,
    objective = "reg:squarederror",  # Regression with squared loss
    verbose = 0,               # Show no running messages
    early_stopping_rounds = 30) # Stop if no improvement for 30 consecutive trees (early stopping)

# Add  min cross-validation RMSE and optimal number of trees for model with lowest cross-validation RMSE to  grid  
hyper_grid_GBRT$optimal_trees[i] = which.min(GBRT_tuning$evaluation_log$test_rmse_mean)
hyper_grid_GBRT$min_RMSE[i] = min(GBRT_tuning$evaluation_log$test_rmse_mean)

print(round((i/nrow(hyper_grid_GBRT)*100),4)) # to check progress of loop in percent 
}


### Fitting model using the optimal tuning parameters ### 

# Optimal hyperparameter w.r.t. tuning procedure 
row_num_opt = which.min(hyper_grid_GBRT$min_RMSE) # row index number in grid object with lowest cross-validation error

params_optimal = list(
  eta = hyper_grid_GBRT[row_num_opt,1],
  max_depth = hyper_grid_GBRT[row_num_opt,2],
  min_child_weight = hyper_grid_GBRT[row_num_opt,3],
  subsample =  hyper_grid_GBRT[row_num_opt,4],
  colsample_bytree =  hyper_grid_GBRT[row_num_opt,5])


set.seed(123) # for reproducibility


GBRT_model = xgboost(
  params = params_optimal,
  data = x_train,
  label = y_train,
  nrounds = hyper_grid_GBRT[row_num_opt,6],
  objective = "reg:squarederror",
  verbose = 0)

### Testing model on in and out-of-sample data ###

regression_GBRT_fitted_train = predict(GBRT_model, newdata = x_train)

regression_GBRT_fitted_test = predict(GBRT_model, newdata = x_test)

RMSE_GBRT_train = RMSE(df_train$price_var,regression_GBRT_fitted_train)
RMSE_GBRT_test = RMSE(df_test$price_var,regression_GBRT_fitted_test)


```


# Traditional regression-based statistical models

The following code chunks correspond to different regression-based models, we apply. More details w.r.t. each are provided in our term paper within the corresponding (sub-)sections.

## Mean (of training data) as estimator

```{r}
# Calculating the mean to use as estimator
mean_train = mean(df_train$price_var)


### Testing model on in and out-of-sample data ###

RMSE_Mean_train = RMSE(df_train$price_var,mean_train)
RMSE_Mean_test = RMSE(df_test$price_var,mean_train)

```


## Linear regression: Pooled OLS

```{r}
### Training and specification tests for Pooled OLS model ### 

# Training linear model
linear_OLS = lm(data=df_train %>% subset(select = -c(year,X)),
                formula= price_var ~ . )

# summary(linear_OLS, robust= TRUE) # Apply robust standard errors to cope with heteroscedastisity

### Specification tests ###

# Test for heteroscestasticity (Breusch-Pagan Test)
bptest(linear_OLS)

# Visualize residuals: Plot Residuals vs. fitted values
plot(fitted(linear_OLS), resid(linear_OLS), xlab='Fitted Values', ylab='Residuals')
abline(0,0)

# Test for autocorrelation (Durban-Watson Test)
dwtest(linear_OLS)

### Testing model on in and out-of-sample data ###

ols_train_fitted = predict(linear_OLS, newdata = df_train)
ols_test_fitted = predict(linear_OLS, newdata = df_test)

RMSE_OLS_train = RMSE(df_test$price_var,ols_train_fitted)
RMSE_OLS_test = RMSE(df_test$price_var,ols_test_fitted)
```


## Linear regression: Pooled Weighted Least Squares (WLS)

```{r}
### Training and specification tests for Pooled Weighted Least Squares model ### 

x_train = data.matrix(subset(df_train, select = -c(year,X,price_var)))


### Step 1: Estimate the (unknown) variances ###
y_train_hat = linear_OLS$fitted # fitted Pooled OLS price_var values

squ_log_residuals = log((df_train$price_var - y_train_hat)^2) # squared log residuals

cond_mean = lm(squ_log_residuals ~ x_train)

var_estimated = exp(cond_mean$fitted)

### Step 2: Calculating the weights of the variance-covariance matrix ###
weights_WLS = 1 / var_estimated

### Step 3: Fitting the WLS model using the estimated variance-covariance matrix ###

linear_WLS = lm(data=df_train %>% subset(select = -c(year,X)),
                formula= price_var ~ . ,
                weights = weights_WLS)

summary(linear_WLS)


### Specification tests ###

# Test for heteroscestasticity (Breusch-Pagan Test)
bptest(linear_WLS)

# Visualize residuals: Plot Residuals vs. fitted values
plot(fitted(linear_WLS), resid(linear_WLS), xlab='Fitted Values', ylab='Residuals')
abline(0,0)

### Testing model on in and out-of-sample data ###
wls_train_fitted = predict(linear_WLS, newdata = df_train)
wls_test_fitted = predict(linear_WLS, newdata = df_test)

RMSE_WLS_train = RMSE(df_train$price_var,wls_train_fitted)
RMSE_WLS_test = RMSE(df_test$price_var,wls_test_fitted)

```


# Mincer-Zarnowitz-Regression of ou-of-sample predictions and statistical tests

We compare the performance of our models among other things by the R-squared of a Mincer-Zarnowitz-Regression which is a regression of the true values on the fitted values received from on applied model at a time. 

```{r}
### Create function to simplify the calculations for the individual t-tests ###
custom_test =function(model){
  t_statistic_constant = (coef(model)[1] - 0)/vcov(model)[1,1]
  p_value_constant = 1 - pnorm(abs(t_statistic_constant))
  t_statistic_coefficient = (coef(model)[2] - 1)/vcov(model)[2,2] 
  p_value_coefficient = 1 - pnorm(abs(t_statistic_coefficient))

  results = matrix(c(t_statistic_constant,t_statistic_coefficient,p_value_constant,p_value_coefficient),
                   nrow=2,ncol=2, byrow = F,
                   dimnames = list(c("constant", "coefficient"),c("t-statistic", "p-value")))
  return(results)
}

### Mincer-Zarnowitz-Regression fo each model and related test ###

MZ_Mean = lm(df_test %>% subset(select = -c(year,X)),
                formula= price_var ~ rep(mean_train,3726))

# OLS
MZ_OLS = lm(df_test %>% subset(select = -c(year,X)),
                formula= price_var ~ ols_test_fitted)

# Test Hypothesis: coef = 1 and intercept = 0 simultaneously (use Wald test)   
wald.test(Sigma = vcov(MZ_OLS), b = coef(MZ_OLS), Terms = c(1,2), H0 = c(0,1))

# Test Hypothesis: coef = 1 and intercept = 0 separately
custom_test(MZ_OLS)


# WLS
MZ_WLS = lm(df_test %>% subset(select = -c(year,X)),
                formula= price_var ~ wls_test_fitted)

# Test Hypothesis: coef = 1 and intercept = 0 (use Wald test)   
wald.test(Sigma = vcov(MZ_WLS), b = coef(MZ_WLS), Terms = c(1,2), H0 = c(0,1))

# Test Hypothesis: coef = 1 and intercept = 0 separately
custom_test(MZ_WLS)


# Ridge
MZ_Ridge = lm(df_test %>% subset(select = -c(year,X)),
                formula= price_var ~ ridge_test_fitted)

# Test Hypothesis: coef = 1 and intercept = 0 (use Wald test)   
wald.test(Sigma = vcov(MZ_Ridge), b = coef(MZ_Ridge), Terms = c(1,2), H0 = c(0,1))

# Test Hypothesis: coef = 1 and intercept = 0 separately
custom_test(MZ_Ridge)


# Lasso
MZ_Lasso = lm(df_test %>% subset(select = -c(year,X)),
                formula= price_var ~ lasso_test_fitted)

# Test Hypothesis: coef = 1 and intercept = 0 (use Wald test)   
wald.test(Sigma = vcov(MZ_Lasso), b = coef(MZ_Lasso), Terms = c(1,2), H0 = c(0,1)) # does not make sense, since lasso collapses to intercept/mean


# Tree
MZ_Tree =  lm(df_test %>% subset(select = -c(year,X)),
                formula= price_var ~ regression_tree_fitted)

# Test Hypothesis: coef = 1 and intercept = 0 (use Wald test)   
wald.test(Sigma = vcov(MZ_Tree), b = coef(MZ_Tree), Terms = c(1,2), H0 = c(0,1)) # does not make sense, since tree model collapses to mean


# Tree utilizing Bagging 
MZ_Bag_Tree = lm(df_test %>% subset(select = -c(year,X)),
                formula= price_var ~ regression_bag_tree_fitted)

# Test Hypothesis: coef = 1 and intercept = 0 (use Wald test)   
wald.test(Sigma = vcov(MZ_Bag_Tree), b = coef(MZ_Bag_Tree), Terms = c(1,2), H0 = c(0,1))

# Test Hypothesis: coef = 1 and intercept = 0 separately
custom_test(MZ_Bag_Tree)


# Random Forest
MZ_Random_Forest = lm(df_test %>% subset(select = -c(year,X)),
                formula= price_var ~ regression_random_forest_fitted_test$predictions)

# Test Hypothesis: coef = 1 and intercept = 0 (use Wald test)   
wald.test(Sigma = vcov(MZ_Random_Forest), b = coef(MZ_Random_Forest), Terms = c(1,2), H0 = c(0,1))

# Test Hypothesis: coef = 1 and intercept = 0 separately
custom_test(MZ_Random_Forest)


# GBRT
MZ_GBRT = lm(df_test %>% subset(select = -c(year,X)),
                formula= price_var ~ regression_GBRT_fitted_test)

# Test Hypothesis: coef = 1 and intercept = 0 (use Wald test)   
wald.test(Sigma = vcov(MZ_GBRT), b = coef(MZ_GBRT), Terms = c(1,2), H0 = c(0,1))

# Test Hypothesis: coef = 1 and intercept = 0 separately
custom_test(MZ_GBRT)


MZ_R2 = as.data.frame(c(summary(MZ_OLS)$r.squared,
                        summary(MZ_WLS)$r.squared,
                        summary(MZ_Lasso)$r.squared,
                        summary(MZ_Ridge)$r.squared,
                        summary(MZ_Random_Forest)$r.squared,
                        summary(MZ_GBRT)$r.squared))

```


# Visualizations 

The following code-chunks are used to create different diagrams and figures for our presentation and term paper. 


## Data Cleaning Process (Sankey diagram)

```{r}
# A: Observation (rows)
nodes = data.frame("name" = 
 c("2014 (3808)", "2015 (4120)", "2016 (4797)", "2017 (4960)","2018 (4392)","Common Stocks (18630)", "Without Outliers (16694)","","Imputation (18630)", "Training data (14904)", "Test data (3726)"))

links = as.data.frame(matrix(c(
 0, 5, 3808, # 2014 - Common Stocks
 1, 5, 4120, # 2015 - Common Stocks
 2, 5, 4797, # 2016 - Common Stocks
 3, 5, 4960, # 2017 - Common Stocks
 4, 5, 4392, # 2018 - Common Stocks
 5, 6, 18630, # Common Stocks - Without Outliers
 6, 7, 16694, # Without Outliers -
 7, 8, 18630,
 8, 9, 14904, # Imputation - Training Data
 8, 10, 3726), # Imputation - Test Data
 byrow = TRUE, ncol = 3))

 names(links) = c("source", "target", "value")

 sankeyNetwork(Links = links, Nodes = nodes,
 Source = "source", Target = "target",
 Value = "value", NodeID = "name",
 units = 'observations', fontSize= 20, nodeWidth = 40, sinksRight = TRUE)
 
 # B: Features (columns)

nodes = data.frame("name" = 
 c("All Feautures (223)", "Adding Sector Dummies (233)", "Columns < 30% of zeros (171)","", "Deleting identical columns (163)", "Columns with to much collinearity (136)"))

links = as.data.frame(matrix(c(
  0, 1, 223,
  1, 2, 233,
  2, 3, 171,
  3, 4, 163,
  4, 5, 136),
 byrow = TRUE, ncol = 3))

 names(links) = c("source", "target", "value")

 sankeyNetwork(Links = links, Nodes = nodes,
 Source = "source", Target = "target",
 Value = "value", NodeID = "name",
 units = 'observations', fontSize= 17, nodeWidth = 40, sinksRight = TRUE)
 
```


## Out-of-sample RMSE accros different models

```{r}

RMSE_train = data.frame(RMSE = c(RMSE_Mean_train,RMSE_OLS_train,RMSE_WLS_train,RMSE_Ridge_train,RMSE_Lasso_train,RMSE_Random_Forest_train,RMSE_GBRT_train),Model = c("Mean","OLS","WLS","Ridge","Lasso","RF","GBRT" ), Sample = rep("In-Sample",7), row = c(3,7,5,2,4,1,6))

RMSE_test = data.frame(RMSE = c(RMSE_Mean_test,RMSE_OLS_test,RMSE_WLS_test,RMSE_Ridge_test,RMSE_Lasso_test,RMSE_Random_Forest_test,RMSE_GBRT_test),Model = c("Mean","OLS","WLS","Ridge","Lasso","RF","GBRT" ),Sample = rep("Out-of-Sample",7), row = c(3,7,5,2,4,1,6))

RMSE_all = rbind(RMSE_train,RMSE_test)

ggplot(data = RMSE_all, aes(x = reorder(Model, row), y = RMSE, fill = Sample)) +
  geom_bar(stat="identity", position = "dodge") +
  labs(x = "", y = "RMSE",fill="") +
  coord_cartesian(ylim=c(50,70)) +
  theme_light(base_size = 14)
```


## Out-of-sample R^{2} from Mincer-Zarnowitz-Regression

```{r}
Rsqu = data.frame(R2 = c(rep(0,7)),
                  Model = c("OLS","WLS","Ridge","Lasso","Tree","Bagging Tree","RF" )) %>% mutate(row = row_number())

# OLS and WLS
Rsqu[1,1] = summary(MZ_OLS)$r.squared * 100
Rsqu[2,1] = summary(MZ_WLS)$r.squared * 100

ggplot(data = Rsqu, aes(x = reorder(Model,row), y = R2)) +
  geom_bar(stat="identity", fill=c("#eb8060","#eb8060","steelblue","steelblue","steelblue","steelblue","steelblue")) +
  labs(x = "", y = TeX(r"($R^{2}_{MZ}*100$)")) +
  geom_text(aes(label=round(R2,3)), vjust=1.6, color="white", size=3.5) +
  theme_light()

# Ridge and Lasso
Rsqu[3,1] = summary(MZ_Ridge)$r.squared * 100
Rsqu[4,1] = summary(MZ_Lasso)$r.squared * 100

ggplot(data = Rsqu, aes(x = reorder(Model,row), y = R2)) +
  geom_bar(stat="identity", fill=c("steelblue","steelblue","#eb8060","#eb8060","steelblue","steelblue","steelblue")) +
  labs(x = "", y = TeX(r"($R^{2}_{MZ} *100$)")) +
  geom_text(aes(label=round(R2,3)), vjust=1.6, color="white", size=3.5) +
  theme_classic()

# Tree 
Rsqu[5,1] = summary(MZ_Tree)$r.squared * 100

ggplot(data = Rsqu, aes(x = reorder(Model,row), y = R2)) +
  geom_bar(stat="identity", fill=c("steelblue","steelblue","steelblue","steelblue","#eb8060","steelblue","steelblue")) +
  labs(x = "", y = TeX(r"($R^{2}_{MZ} *100$)")) +
  geom_text(aes(label=round(R2,3)), vjust=1.6, color="white", size=3.5) +
  theme_classic()

# Tree Bagging
Rsqu[6,1] = summary(MZ_Bag_Tree)$r.squared * 100

ggplot(data = Rsqu, aes(x = reorder(Model,row), y = R2)) +
  geom_bar(stat="identity", fill=c("steelblue","steelblue","steelblue","steelblue","steelblue","#eb8060","steelblue")) +
  labs(x = "", y = TeX(r"($R^{2}_{MZ} *100$)")) +
  geom_text(aes(label=round(R2,3)), vjust=1.6, color="white", size=3.5) +
  theme_classic()

# Random Forest
Rsqu[7,1] = summary(MZ_Random_Forest)$r.squared * 100

ggplot(data = Rsqu, aes(x = reorder(Model,row), y = R2)) +
  geom_bar(stat="identity", fill=c("steelblue","steelblue","steelblue","steelblue","steelblue","steelblue","#eb8060")) +
  labs(x = "", y = TeX(r"($R^{2}_{MZ} *100$)")) +
  geom_text(aes(label=round(R2,3)), vjust=1.6, color="white", size=3.5) +
  theme_classic()



### Visualization for our term paper including result for GBRT ###
Rsqu = data.frame(R2 = MZ_R2[,1] * 100, # R2 in percentage points
                  Model = c("OLS","WLS","Lasso","Ridge","RF","GBRT"), row= c(2,5,1,3,6,4))


ggplot(data = Rsqu, aes(x = reorder(Model,row), y = R2)) +
  geom_bar(stat="identity", fill=c("steelblue","steelblue","steelblue","steelblue","steelblue","steelblue")) +
  labs(x = "", y = TeX(r"($R^{2}_{MZ} *100$)")) +
  geom_text(aes(label=round(R2,3)), vjust=1.6, color="white", size=3.5) +
  theme_classic(base_size = 14)

```

## Scatter plot for tree example

This scatter plot is used to explain the basic algorithm of regression trees. 

```{r}
# Example Tree model

ggplot(data=df_train, aes(x= Days.Payables.Outstanding, y=EPS)) +
  geom_point() +
  geom_segment(aes(x = -500 , y = 0, xend = 33000, yend = 0,color = "#eb8060")) +
  geom_vline(xintercept = 33000, color="#eb8060") +
  annotate("text", x = 40000, y = -7500, color = "#eb8060", label = "R1", size=6) +
  annotate("text", x = 15000, y = 1000, color = "#eb8060", label = "R2", size=6) +
  annotate("text", x = 15000, y = -7500, color = "#eb8060", label = "R3", size=6) +
  labs(x="Days Payable Outstanding",y="Earnings Per Share (EPS)") +
  theme_classic(base_size = 16) +
  theme(legend.position = "none")
```

## Tree model figrue

This figure of a Regression Tree model helps to explain the structure of Tree models. 

```{r}
tree_example = rpart(
    formula = price_var ~ .,
    data    = df_train %>% subset(select = -c(year,X)),
    method  = "anova",
    control = list( minsplit = 5, # min obs. in a nod to perform split
                    minbucket = round(5), # min number of obs. in terminal leaf
                    xval = 10, # number of folds in cross-validation
                    cp = 0.03139987, # complexity parameter
                    maxdepth = 2)) # max depth of any node of the final tree)

# Visualize Tree
rpart.plot(tree_example , type = 2)

# Test model on out-of-sample test data
regression_tree_fitted_tree_example = predict(tree_example,
                                      newdata = df_test %>% subset(select = -c(year,X)) )

MSE_Tree = mean((df_test$price_var - regression_tree_fitted_tree_example)^2)

RMSE_Tree = RMSE(df_test$price_var,regression_tree_fitted_tree_example)

```


## Example figure GBRT model

```{r}

# OPtimal hyperparameters w.r.t. our tuning procress

set.seed(123)

GBRT_example = xgb.cv(
  params = params_optimal,
  data = x_train,
  label = y_train,nfold = 5,
  nrounds = 1000, # include more than optimal number of iterations/trees to visualize overfitting
  objective = "reg:squarederror",
  verbose = 0)

colors = c("training error" ="steelblue","validation error"="#eb8060")

# Visualize training and validation MSE
ggplot(data = GBRT_example$evaluation_log) +
  geom_line(aes(iter, train_rmse_mean, color = "training error" ), size=1) + 
  geom_line(aes(iter, test_rmse_mean, color = "validation error"), size=1) +
  geom_vline(xintercept = as.numeric(GBRT_example$evaluation_log[which.min(GBRT_example$evaluation_log$test_rmse_mean),1]), color="black",lty = "dashed",size=0.7)+  # number of tree with lowest RMSE
  labs(x="Number of Trees/ Iterations",y="RMSE", color="") + 
  scale_x_continuous(breaks=seq(0,1000,100)) +
  theme_light()

```

